---
title: "first-analysis"
output: html_document
date: "2024-02-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is how I felt completing this website! Anyways lets get into the results shall we

```{r echo=FALSE, warning=FALSE, results='hide'}
# Load required packages
library(grid)

# Set the working directory to where your image is located
setwd("C://Users//Jmons//OneDrive//Documents//R-Class work")

library(jpeg)
# Read the image file
img <- readJPEG("Funny_R_Meme.jpeg") # You can also use readJPEG() for JPEG files or readTIFF() for TIFF files

# Display the image
grid::grid.raster(img)

```





# The Top 20 

<span style="color:rgb(255, 20, 147);"> There are so many Species of the genus Candida so just for fun, before we look at the truly pathogenic ones lets see how about 20 of them reacted in growing in different sugars 

<span style="color:rgb(255, 20, 147);"> **Here is a heat map looking at how much of the sugars were used based on each species**


```{r echo=FALSE, warning=FALSE, results='hide'}

setwd("C:/Users/Jmons/OneDrive/Documents/R-Class work/MyCharts")
getwd()

library(readr)
Shorted_C_NormGR_OnlySugars <- read.csv("C:/Users/Jmons/OneDrive/Documents/R-Class work/MyCharts/data/Shorted_C_NormGR_OnlySugars.csv")

View(Shorted_C_NormGR_OnlySugars)

library(dplyr)
Candida_data <- Shorted_C_NormGR_OnlySugars %>% filter(Genus == "Candida")

# View the filtered data
View(Candida_data)
# or print the first few rows in the console
head(Candida_data)

#making violent plot

# Load necessary libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)

# Load and prepare the data
Shorted_C_NormGR_OnlySugars <- read.csv("C:/Users/Jmons/OneDrive/Documents/R-Class work/MyCharts/data/Shorted_C_NormGR_OnlySugars.csv")
Candida_data <- Shorted_C_NormGR_OnlySugars %>% filter(Genus == "Candida")

Candida_long <- Candida_data %>%
  pivot_longer(
    cols = -c(PU, Species, Genus, NRRL, CBS, Other, Hittinger), # Adapt these column names as needed
    names_to = "Sugar",
    values_to = "Measurement"
  )





```


```{r echo=FALSE, warning=FALSE, results='hide'}
##########Table
# Load necessary libraries
library(readr)

# Read the data
Shorted_C_NormGR_OnlySugars <- read.csv("C:/Users/Jmons/OneDrive/Documents/R-Class work/MyCharts/data/Shorted_C_NormGR_OnlySugars.csv")

# Print all column names to check their exact spelling
print(names(Shorted_C_NormGR_OnlySugars))

# Assuming 'library(dplyr)' has already been loaded
# Filter data for the genus "Candida"
Candida_data <- Shorted_C_NormGR_OnlySugars %>% 
  filter(Genus == "Candida")

# Adjust this select statement to match the confirmed column names
Candida_data_selected <- Candida_data %>%
  select(Species, everything())  # This line is just a placeholder; adjust based on actual column names
# Print the first few rows to confirm the structure
head(Candida_data_selected)

# Optionally, use View to inspect the data frame in a more interactive way
View(Candida_data_selected)

# Assuming you've already loaded the necessary libraries and have the Candida_data_selected dataframe ready
library(dplyr)

# Exclude columns 2 through 7 from the dataframe
Candida_data_selected <- Candida_data_selected %>%
  select(-c(2:7))

# Print the first few rows to confirm that the columns have been excluded
head(Candida_data_selected)

# Optionally, use View to inspect the data frame in a more interactive way
View(Candida_data_selected)



```


```{r echo=FALSE, fig.height=5, fig.width=8, warning=FALSE, results='hide'}
# Load necessary libraries
library(ggplot2)
library(tidyr)
library(dplyr)

# Assuming Candida_data_selected is already loaded
# Identify the top 20 unique species
top_species <- unique(Candida_data_selected$Species)[1:20]

# Filter the data to include only the top 20 species
filtered_data <- Candida_data_selected %>%
  filter(Species %in% top_species)

# Pivot the data to long format
long_data <- pivot_longer(
  filtered_data, 
  cols = -Species,  # Keep the Species column as the identifier
  names_to = "SugarType", 
  values_to = "Measurement"
)
# Create the heatmap
heatmap <- ggplot(long_data, aes(x = SugarType, y = reorder(Species, desc(Species)), fill = Measurement)) +
  geom_tile() +  # Creates the heatmap tiles
  scale_fill_gradient(
    low = "pink", 
    high = "red", 
    limits = c(0, 7), 
    breaks = seq(0, 7, by = 1),
    guide = guide_colourbar(
      barheight = unit(10, "cm"), # Customize the height of the color bar
      barwidth = unit(1, "cm"),  # Customize the width of the color bar
      title.position = "top",
      title.hjust = 0.5,
      label.position = "right", # Correct label position for vertical direction
      label.hjust = 0.5
    )
  ) +
  labs(title = "Heatmap of Sugar Measurements by Species (Top 20)",
       x = "Type of Sugar", 
       y = "Species") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),  # Rotate x-axis labels for readability
        axis.text.y = element_text(angle = 0, hjust = 1),  # Ensure y-axis labels are readable
        legend.position = "right",
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 8))

# Print the heatmap
print(heatmap)


```

<span style="color:rgb(255, 20, 147);">Pretty cool right? Based on these results we see that relatively the same amount of sugar is taken up by each species but it appears that **Mannose and Fluctose** tend to be the favorite if they got to choose

# Top 100 Analysis

<span style="color:rgb(255, 20, 147);">Now lets expand this a bit and look at 100 different species on a **heatmap**. We still see this commonality of the yeast liking Fructose and Mannose the most

```{r echo=FALSE, fig.height=30, fig.width=20, warning=FALSE, results='hide'}
# Assuming Candida_data_selected is already loaded
# Identify the top 100 unique species
top_species <- unique(Candida_data_selected$Species)[1:100]

# Filter the data to include only the top 100 species
filtered_data <- Candida_data_selected %>%
  filter(Species %in% top_species)

# Pivot the data to long format
long_data <- pivot_longer(
  filtered_data, 
  cols = -Species,  # Keep the Species column as the identifier
  names_to = "SugarType", 
  values_to = "Measurement"
)
# Create the heatmap with adjusted settings for a taller display
# Load necessary libraries
library(ggplot2)
library(tidyr)
library(dplyr)

# Assuming long_data is ready and contains your heatmap data
# Adjust the heatmap settings for a taller display
# Assuming the necessary libraries and data have been loaded
heatmap <- ggplot(long_data, aes(x = SugarType, y = reorder(Species, desc(Species)), fill = Measurement)) +
  geom_tile() +  # Creates the heatmap tiles
  scale_fill_gradient(
    low = "darkred", 
    high = "pink", 
    limits = c(0, 7), 
    breaks = seq(0, 7, by = 1),
    guide = guide_colourbar(
      barheight = unit(10, "cm"), # Customize the height of the color bar
      barwidth = unit(1, "cm"),  # Customize the width of the color bar
      title.position = "top",
      title.hjust = 0.5,
      label.position = "right", # Correct label position for vertical direction
      label.hjust = 0.5
    )
  ) +
  labs(title = "Heatmap of Sugar Measurements by Top 100 Species",
       x = "Type of Sugar", 
       y = "Species") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 30),  # Rotate and adjust font size for X-axis labels
        axis.text.y = element_text(size = 20, angle = 0, vjust = 0.5),  # Smaller font size for Y-axis labels
        plot.title = element_text(size = 25),
        axis.title = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 16), # Decreased font size
        legend.position = "right",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Save the heatmap with adjusted legend settings
ggsave("tall_heatmap.png", plot = heatmap, width = 10, height = 50, units = "in", dpi = 300, limitsize = FALSE)

print(heatmap)


```


<span style="color:rgb(255, 20, 147);">Another way to look at this analysis is through the use of a violin plot, showing how much is being taken up by the yeast based on the sugar type


```{r echo=FALSE, fig.height=20, fig.width=40, warning=FALSE, results='hide'}
# Load necessary libraries
library(ggplot2)
library(tidyr)
library(dplyr)

# Assuming Candida_data_selected is already loaded
# Identify the top 100 unique species
top_species <- unique(Candida_data_selected$Species)[1:100]

# Filter the data to include only the top 100 species
filtered_data <- Candida_data_selected %>%
  filter(Species %in% top_species)

# Pivot the data to long format
long_data <- pivot_longer(
  filtered_data, 
  cols = -Species,  # Keep the Species column as the identifier
  names_to = "SugarType", 
  values_to = "Measurement"
)

# Create the violin plot with increased width for each violin
violin_plot <- ggplot(long_data, aes(x = SugarType, y = Measurement, fill = SugarType)) +
  geom_violin(trim = TRUE, adjust = 1, width = 1) + ylim(0, 2) +  # Adjust violin width
  labs(title = "Violin Plot of Sugar Measurements by Top 100 Species",
       x = "Type of Sugar",
       y = "Measurement") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 40),  # Adjust the size and angle of x-axis labels
        axis.text.y = element_text(size = 40),  # Increase the size of y-axis labels
        plot.title = element_text(size = 40),
        axis.title = element_text(size = 30),
        legend.title = element_text(size = 30),
        legend.text = element_text(size = 25),  # Adjusted font size
        legend.position = "right",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Save the violin plot with increased width and readability
ggsave("violin_plot_adjusted_wide.png", plot = violin_plot, width = 40, height = 20, units = "in", dpi = 300, limitsize = FALSE)

# Print the violin plot
print(violin_plot)



```


# The Most Pathogenic Yeast


<span style="color:rgb(255, 20, 147);">Now lets look at the top dogs, that have been known for being the most pathogenic yeast in the worlds, those these isn't a lot, this many is still concerting.

```{r echo=FALSE, warning=FALSE, results='hide'}
setwd("C:/Users/Jmons/OneDrive/Documents/R-Class work/MyCharts")
getwd()


# Load necessary library
library(readr)

# Read the data from the CSV file
data <- read_csv("C:/Users/Jmons/OneDrive/Documents/R-Class work/MyCharts/data/NormGR_OnlySugars.csv")
# Define the vector of species names
species_to_include <- c("pichia kudriavzevii", "Meyerozyma guilliermondii", 
                        "Candida parapsilosis", "Candida orthopsilosis", 
                        "Candida dubliniensis", "Candida albicans", 
                        "Candida tropicalis", "Clavispora lusitaniae", 
                        "Candida auris", "Nakaseomyces delphensis", 
                        "Candida metapsilosis")

# Filter the data to include only these species
ToxicYeast_data <- data[data$Species %in% species_to_include, ]

# Remove columns by index
ToxicYeast_data <- ToxicYeast_data[ , -c(1, 3, 4, 5, 6, 7)]

# Inspect the first few rows of the modified data
head(ToxicYeast_data)


View(ToxicYeast_data)

```
```{r echo=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(tidyr)  # For pivoting data if necessary
library(dplyr)
# Example of pivoting, adjust column names as needed
long_data <- ToxicYeast_data %>%
  pivot_longer(
    cols = -Species,  # Exclude the Species column from pivoting
    names_to = "Sugar",
    values_to = "Measurement"
  )
# Creating a bar plot with ggplot2
# Creating a bar plot with ggplot2, with adjusted bar width
ggplot(long_data, aes(x = Sugar, y = Measurement, fill = Species)) +
  geom_bar(stat = "summary", fun = "mean", position = position_dodge(width = 0.8), width = 0.9) +  # Increased width parameter
  theme_minimal() +
  labs(title = "Average Sugar Measurements by Species",
       x = "Type of Sugar",
       y = "Average Measurement") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
# Save the updated plot with sufficient dimensions to accommodate wider bars
ggsave("sugar_measurements_plot.png", plot = last_plot(), width = 12, height = 8, units = "in")


```
<span style="color:rgb(255, 20, 147);">This is just one way of looking at how these yeast enjoy certain sugars over the other

<span style="color:rgb(255, 20, 147);">Another way to look at this is through the use of a **bee swarm map**, giving a better visual of each value. 

```{r echo=FALSE, warning=FALSE, results='hide'}
# Install ggbeeswarm package if it's not already installed
if (!requireNamespace("ggbeeswarm", quietly = TRUE)) {
  install.packages("ggbeeswarm")
}

# Install the tidyverse package if not already installed
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}

# Load necessary libraries
library(readr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)

# Pivot the data to long format where each row is a single measurement for a type of sugar
ToxicYeast_data_long <- pivot_longer(ToxicYeast_data,
                                    cols = -Species,  # Exclude the species column from pivoting
                                    names_to = "SugarType",  # This will hold the names of the sugar columns
                                    values_to = "Measurements"  # This will hold the measurement values
                                    )

# Create the bee swarm plot
ggplot(ToxicYeast_data_long, aes(x = Species, y = Measurements, color = SugarType)) +
  geom_beeswarm() +
  theme_minimal() +
  labs(title = "Bee Swarm Plot of Measurements by Species",
       x = "Species",
       y = "Measurements",
       color = "Type of Sugar") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Improve x-axis label readability


```
